---
title: "R Notebook"
output: html_notebook
---
 
# Time Series Analysis on Microsoft Stock
```{r}

MSFT_day=read.csv("https://raw.githubusercontent.com/katariyj/Time-Series-Forecasting-S-P-500/master/TestMSFT20051111.csv",header=T)            
head(MSFT_day)
tail(sp500daily)

```
BASIC EDA 
Head() - This will return the first n observations of the dataset. Default is 6
Tail() - This will return the last n observations of the dataset. Default is 6
Summary() - This will return the summary of the statistics like min, median, max, quartiles for each column of the dataset. Count returns missing value if any.
Dim() - This will return the dimensions of the data frame like the number of rows and number of columns in the dataset. The output is a vector.
Hist() - This will return a bar plot in which each bar represents the frequency of cases for a range of values
Str() - returns the classes of each of the columns whether it is numerical or categorical data.
Names() - This will return the column headers.

BASIC EDA OF MSFT

```{r}
head(MSFT_day)

tail(MSFT_day)

summary(MSFT_day)

dim(MSFT_day)

hist(MSFT_day$close)

str(MSFT_day)

names(MSFT_day)

plot(MSFT_day$close, type="l", x = MSFT_day$Time, col="blue")

# Advance EDA
qplot(MSFT_day$close, fill = I("green"), alpha = I(.5), col= I("red"),xlab = "Close",  main="Histogram of MSFT Closing Price", bins = 15)


msftminutes = ts(MSFT_day$close, frequency = 48)

autoplot(stl(msftminutes, s.window = "periodic"), ts.colour = "blue")

ggtsdisplay(MSFT_day$close, main= "MSFT Trend, ACF, PACF Plot")
```

Advance EDA For Minutes Returns for MSFT 

```{r}
MSFT_price = MSFT_day$close
MSFTreturns =diff(MSFT_price)/MSFT_price[-length(MSFT_price)]


qplot(MSFTreturns, fill = I("blue"), alpha = I(.5), col= I("red"),xlab = "MSFT Returns",  main="Histogram of MSFT Returns", bins = 30)


MSFT_returns = ts(MSFTreturns, frequency = 48)

autoplot(stl(MSFT_returns, s.window = 'periodic'), ts.colour = 'blue')

ggtsdisplay(MSFTreturns, main = "MSFT Returns Trend, ACF, PACF Plot")


```



Transformations and EDA - LOG Returns of MSFT
```{r}
MSFT_logreturns1=diff(log(MSFT_day$close))


qplot(MSFT_logreturns1, fill = I("blue"), alpha = I(.5), col= I("red"),xlab = "MSFT Log Returns",  main="Histogram of MSFT Log Returns", bins = 30)


MSFT_logreturns2 = ts(MSFT_logreturns1, frequency = 48)

autoplot(stl(MSFT_logreturns2, s.window = 'periodic'), ts.colour = 'blue')

ggtsdisplay(MSFT_logreturns1, main = "MSFT Log Returns Trend, ACF, PACF Plot")
```

Compare Histogram of minute returns and Log returns 
```{r}

require(gridExtra)
Minute_Returns = qplot(MSFTreturns, fill = I("green"), alpha = I(.5), col= I("red"),xlab = "MSFT Returns",  main="Histogram of MSFT Returns", bins = 10)

Log_Returns=qplot(MSFT_logreturns1, fill = I("blue"), alpha = I(.5), col= I("red"),xlab = "MSFT Log Returns",  main="Histogram of MSFT Log Returns", bins = 10)
grid.arrange(Minute_Returns, Log_Returns, ncol=2)

```

Check Stats of the data 
```{r}

ggplot(data.frame(MSFT_logreturns1), aes(sample=MSFT_logreturns1))+stat_qq()

t.test(MSFT_logreturns1)

library(fBasics) ## This loads the package fBasic
basicStats(MSFTreturns)
basicStats(MSFT_logreturns1)

ggplot(data.frame(MSFTreturns), aes(sample=MSFTreturns))+stat_qq()
t.test(MSFTreturns)

## Normality tests

normalTest(MSFTreturns,method="jb")

autoplot(acf(MSFTreturns,lag=15, plot = FALSE))

#Independence test
Box.test(MSFTreturns,lag=10)

Box.test(MSFTreturns,lag=10,type="Ljung")
```

Model Building

```{r}
model1=ar(MSFTreturns,method='mle')
model1
model1_residuals = model1$resid[!is.na(model1$resid)]

ggtsdisplay(model1_residuals, main = "Model 1 residuals Trend, ACF, PACF Plot")

#p value is greater than alpha so residuals are independent
Box.test(model1_residuals,lag=10,type='Ljung') 

```
Other models
```{r}
model2=arima(MSFTreturns,order=c(4,0,3))
model2

ggtsdisplay(model2$residuals, main = "Model 2 residuals Trend, ACF, PACF Plot")

#p value greater than alpha so the residuals are independent
Box.test(model2$residuals,lag=10,type='Ljung') 


model3=arima(MSFTreturns,order=c(3,0,4))
model3

ggtsdisplay(model3$residuals, main = "Model 3 residuals Trend, ACF, PACF Plot")

#p value greater than alpha so the residuals are independent
Box.test(model3$residuals,lag=10,type='Ljung') 


#p value greater than alpha so the residuals are independent
model4=arima(MSFTreturns,order=c(3,0,3))
model4

ggtsdisplay(model4$residuals, main = "Model 4 residuals Trend, ACF, PACF Plot")

Box.test(model4$residuals,lag=10,type='Ljung') 


model5=arima(MSFTreturns,order=c(6,0,6))
model5

ggtsdisplay(model5$residuals, main = "Model 2 residuals Trend, ACF, PACF Plot")

Box.test(model5$residuals,lag=10,type='Ljung') 
```

We select Model 4
```{r}

ggtsdiag(model4)

```



```{r}
poly1=c(1,-model4$coef[1:3])
roots=polyroot(poly1)
roots

Mod(roots)
```

```{r}
prediction = predict(model4, 10)
prediction
```