---
title: "R Notebook"
output: html_notebook
---
 
Set the working directory and read the data from the respective folder
```{r}
sp500daily=read.csv("https://raw.githubusercontent.com/NikhilKorati/Time-Series/master/GSPCdaily26Feb2018.csv",header=T)        
head(sp500daily)
tail(sp500daily)

```
BASIC EDA 
Head() - This will return the first n observations of the dataset. Default is 6
Tail() - This will return the last n observations of the dataset. Default is 6
Summary() - This will return the summary of the statistics like min, median, max, quartiles for each column of the dataset. Count returns missing value if any.
Dim() - This will return the dimensions of the data frame like the number of rows and number of columns in the dataset. The output is a vector.
Hist() - This will return a bar plot in which each bar represents the frequency of cases for a range of values
Str() - returns the classes of each of the columns whether it is numerical or categorical data.
Names() - This will return the column headers.

BASIC EDA OF SP500
```{r}
head(sp500daily)

tail(sp500daily)

summary(sp500daily)

dim(sp500daily)

hist(sp500daily$Adj.Close)

str(sp500daily)

names(sp500daily)

plot(sp500daily$Adj.Close, type = "l",col="red")


# Advance EDA
library(ggplot2)

qplot(sp500daily$Adj.Close, fill = I("blue"), alpha = I(.5), col= I("red"),xlab = "Adj Close",  main="Histogram of SP500 Adj Closing Price", bins = 30)


install.packages("ggfortify")

library(ggfortify)

library(stats)
library(forecast)

sp500close = ts(sp500daily$Adj.Close, frequency = 48)

autoplot(stl(sp500close, s.window = 'periodic'), ts.colour = 'blue')

ggtsdisplay(sp500daily$Adj.Close, main = "SP500 Trend, ACF, PACF Plot")


```

Advance EDA For Daily Returns for SP500

```{r}
sp500_price = sp500daily$Adj.Close
sp500returns =diff(sp500_price)/sp500_price[-length(sp500_price)]


qplot(sp500returns, fill = I("blue"), alpha = I(.5), col= I("red"),xlab = "SP500 Returns",  main="Histogram of SP500 Returns", bins = 30)


sp500_returns = ts(sp500returns, frequency = 48)

autoplot(stl(sp500_returns, s.window = 'periodic'), ts.colour = 'blue')

ggtsdisplay(sp500returns, main = "SP500 Returns Trend, ACF, PACF Plot")

```

Transformations and EDA - LOG Returns of SP500
```{r}
sp500_logreturns1=diff(log(sp500_price))


qplot(sp500_logreturns1, fill = I("blue"), alpha = I(.5), col= I("red"),xlab = "SP500 Log Returns",  main="Histogram of SP500 Log Returns", bins = 30)


sp500_logreturns2 = ts(sp500_logreturns1, frequency = 48)

autoplot(stl(sp500_logreturns2, s.window = 'periodic'), ts.colour = 'blue')

ggtsdisplay(sp500_logreturns1, main = "SP500 Log Returns Trend, ACF, PACF Plot")
```

Compare Histogram of daily returns and Log returns 
```{r}

require(gridExtra)
Daily_Returns = qplot(sp500returns, fill = I("green"), alpha = I(.5), col= I("red"),xlab = "SP500 Returns",  main="Histogram of SP500 Returns", bins = 30)

Log_Returns=qplot(sp500_logreturns1, fill = I("blue"), alpha = I(.5), col= I("red"),xlab = "SP500 Log Returns",  main="Histogram of SP500 Log Returns", bins = 30)
grid.arrange(Daily_Returns, Log_Returns, ncol=2)

```

Check Stats of the data 
```{r}

ggplot(data.frame(sp500_logreturns1), aes(sample=sp500_logreturns1))+stat_qq()
library(fBasics) ## This loads the package fBasic
basicStats(sp500returns)
basicStats(sp500_logreturns1)

t.test(sp500_logreturns1)

ggplot(data.frame(sp500returns), aes(sample=sp500returns))+stat_qq()

t.test(sp500returns)

## Normality tests

normalTest(sp500returns,method="jb")

autoplot(acf(sp500returns,lag=15, plot = FALSE)) # Obtain the ACF plot

##Independence tests
Box.test(sp500returns,lag=10)

Box.test(sp500returns,lag=10,type="Ljung")
```

Model1 - AR Model

```{r}
model1 = ar(sp500returns, method = 'mle')
model1
ggtsdisplay(model1$resid, main= "Residual Plot, ACF, PACF")
```

Model 2


```{r}
model2=arima(x=sp500returns,order=c(0,0,1))
model2
Box.test(model2$resid,lag=10,type='Ljung') 
autoplot(acf(model2$resid, plot = FALSE))

model3=arima(x=sp500returns,order=c(1,1,1))
model3
Box.test(model3$resid,lag=10,type='Ljung') 
autoplot(acf(model3$resid, plot = FALSE))

model4=arima(x=sp500returns,order=c(1,0,1))
model4
Box.test(model4$resid,lag=10,type='Ljung') 
autoplot(acf(model4$resid, plot = FALSE))

model5=arima(x=sp500returns,order=c(2,1,2))
model5
Box.test(model5$resid,lag=10,type='Ljung') 
autoplot(acf(model5$resid, plot = FALSE))

model6=arima(x=sp500returns,order=c(3,0,2))
model6
Box.test(model6$resid,lag=10,type='Ljung') 
autoplot(acf(model6$resid, plot = FALSE))

model7=arima(x=sp500returns,order=c(3,0,3))
model7
Box.test(model7$resid,lag=10,type='Ljung') 
autoplot(acf(model7$resid, plot = FALSE))
```
We select Model 4 


```{r}
ggtsdiag(model4)
```

```{r}
poly1=c(1,-model4$coef[1:3])
roots=polyroot(poly1)
roots

Mod(roots)
```

```{r}
prediction = predict(model4, 10)
prediction

```

